---
title: "loops_and_functions"
author: "Javier Patron"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Task 1

Create two sequences, one called `weekdays` that contains days of the week (\"Monday\", \"Tuesday\", \"Wednesday\", etc.) and one called `transects` that contains the series of transect names \"Transect A\", \"Transect B,\"Transect C\".

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
transect <- c("Transect A", "Transect B", "Transect C")

empty_matrix <- matrix(nrow = length(weekdays), ncol=length(transect))


for (i in seq_along(transect)) {
  for (j in seq_along(weekdays)) {
    mix <- paste(weekdays[j], transect[i])
    
    empty_matrix[j,i] <- mix
  }
}

empty_matrix

```

## Part 2

### Task 3

Write a function called `force` that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in ms2ms2 (recall: F=maF=ma), and returns a statement \"The resulting force is \_\_\_ Newtons.\"

```{r}

force <- function(m, a) {
  F = m*a
  print(paste("The resulting force is", F, "Newtons"))
}

# Test your function

force(2,3)
```

### Task 3

| sci_name            | common_name     |  a_est | b_est |
|:--------------------|:----------------|-------:|------:|
| Chanos chanos       | Milkfish        | 0.0905 |  2.52 |
| Sphyraena barracuda | Great barracuda | 0.0181 |  3.27 |
| Caranx ignobilis    | Giant trevally  | 0.0353 |  3.05 |

W=aLbW=aLb,

```{r}
# QUESTION: Why it does run? 

fish_parns2 <- data.frame(sci_name= c("Chanos chanos", "Sphyraena barracuda", "Caranx ignobilis"), common_name=c("Milfish", "Great barracuda", "Giant trevally"), a_est=c(0.0905, 0.0181, 0.0353), b_est=c(2.52, 3.27, 3.05))
```

Then, write a function called `fish_weight` that allows a user to only enter the *common name* (argument `fish_name`) and *total length* (argument `tot_length`) (in centimeters) of a fish, to return the expected fish weight in grams. Test it out for different species and lengths.

```{r}

fish_parns <- tribble(
  ~sci_name, ~common_name, ~a_est, ~b_est,
  "Chanos chanos", "Milfish", 0.0905, 2.52,
  "Sphyraena Barracuda", "Great barracuda", 0.0181, 3.27,
  "Caranx ignobilis", "Giant trevally", 0.0353, 3.05,
)


fish_weight <- function (name, length) {
 filter_by_name <- filter(fish_parns, common_name == name)
 tot_weight <- (filter_by_name$a_est * length * filter_by_name$b_est)
 print(tot_weight)
}

fish_weight("Milfish", 50)

 

```

Now, try creating a vector of lengths (e.g. 0 to 100, by increments of 1) and ensuring that your function will calculate the fish weight over a *range* of lengths for the given species (try this for milkfish, storing the output weights as `milkfish_weights`.

```{r}

vec <- c(0:100)
fish_weight(a=0.0905,b= 2.52,l=vec)


```

### **Task 4**

Wave power (more accurately wave energy flux) in deep water is approximated by:

$Pdeep=0.5H2T$

where PP is power in kWmkWm (potential power per wave meter), HH is wave height in meters (more specifically, the *significant wave height*), and TT is the wave period in seconds. Learn more [**here**](https://en.wikipedia.org/wiki/Wave_power).

Write a function called `wave_power` that calculates potential ocean wave power given inputs of wave height and period.

Use your `wave_power` function to approximate wave power for a period of 8 seconds, over a range of wave heights from 0 to 3 meters by increments of 0.2 meters.

```{r}
wave_power <- function(h, t) {
  p_deep = 0.5 * h^2 * t
  print(paste("The Potentional Ocean Wave Power is", p_deep, "in kWm"))
}


h <- seq(0,3,by=0.2)
wave_power(t=8, h=h)
```

### **Task 5 (OPTIONAL)**

The wave energy equation changes based on ocean depth. Along the coast of Brenville, which has a very sharp shelf as the wave approaches the coast, wave energy is approximated using the deep ocean equation (the one you used above) for depths \> 12 meters, and a shallow equation for depths \<= 12 meters. The Brenville team estimates shallow wave power by:

Pshallow=0.81H2TPshallow=0.81H2T

Create a function that requires inputs of water depth, wave height and period, then returns the approximated wave power using the correct equation for the depth entered. It should also include a message (hint: use `message()` just like you would use warning!) that lets the user know if the shallow or deep water equation was used.
